Seismic Hazard Estimation of Northern Iran Using Smoothed Seismicity

### 2016-03-13 seismic-catalog.tex
updates to seismic catalog section

\begin{figure*}[ht!] 
	\centering
	\includegraphics[width=0.9\textwidth]{figures/pdf/figure-01}
	\caption{Region of interest and Iranian seismic zones. At the top, the map of Iran and surrounding countries with shaded relief in the background and the Iranian seismic zones \citep[after][]{Karimiparidari2013} and region of interest in the foreground. The dashed box indicates the complete region of interest and the thicker line box indicates the effective area for which seismic hazard is estimated. The effective region of interest is shown in greater detail at the bottom. Details include the location of the major urban centers and seismic faults in northern Iran.}
	\label{fig:iran}
\end{figure*}

\begin{figure*}[ht!] 
	\centering
	\includegraphics[width=0.9\textwidth]{figures/pdf/figure-02}
	\caption{Notable earthquakes in the region of interest and within three major seismic zones in northern Iran. Epicentral locations are indicated with stars. The labels next to each star or linked with a line to avoid congestion show the magnitude and year of the earthquake.}
	\label{fig:selected}
\end{figure*}

\section{Region of Interest}

Seismic activity in the Iranian plateau is dominated by the tectonic contacts between the Arabian, Eurasian, and Indian plates, which confine the region laterally between the Indian and Eurasian plates, and the Arabian and eastern Asia-Minor plates to the east and west, respectively. This setting gives rise to the Zagros, Alborz, and east Iranian mountain ranges, and defines the geology of the region. Based on these conditions, Iranian seismicity is typically divided into various seismic zones, varying from as few as four to as many as twenty-three \citep[e.g.,][]{Stocklin1968, Berberian1976, Tavakoli1999}. Here, we adopt the model proposed by \citet{Karimiparidari2013}, which consists of six seismic zones: Azerbaijan, Alborz, Kopeh-Dagh, Zagros, Central-East, and Makram (Fig.~\ref{fig:iran}).

Within this context, we define our region of interest between longitudes 43.5\textdegree{}E and 61.5\textdegree{}E, and latitudes 34\textdegree{}N and 40\textdegree{}N, as shown in Fig.~\ref{fig:iran}. These limits correspond to the effective region of interest for which we will present results in subsequent sections. Our calculations, however, included data from a slightly larger region with a 0.5\textdegree{}-perimeter as indicated in the figure. This is done to avoid the artificial decay effect that results from smoothing the seismicity along and near the edges of the region of interest. The effective region has an area of 1,703 km $\times$ 667 km, and as shown in Fig.~\ref{fig:iran}, it encloses the northern part of Iran with most of the Azerbaijan seismic zone, the complete zones of Alborz and Kopeh-Dagh, and the northernmost portions of the Zagros and Central-East seismic zones.

At the northeast of Iran, the seismic zone of Azerbaijan is strongly controlled by the North Tabriz Fault (NTF) system. This system is a complex northwest-trending structure with a predominant right-lateral, str\-ike-slip faulting mechanism \citep{Berberian1999}. Historical accounts document the occurrence of earthquakes in this area as far back as to the 9th century, with events in the years 858 ($M$ 6.0), 1042 ($M$ 7.3), and 1304 ($M$ 6.7) \citep{Berberian1999}. In the 18th century, in particular, the NTF ruptured three times in a span period of 65 years, which included the $M$ 7.7 1721 Tabriz earthquake in the southeastern section, the $M$ 7.6 1780 Tabriz earthquake in the northwestern section, and the $M$ 6.3 1786 Marand-Mishu on the Mishu reverse fault and the Sufian segment of the NTF. According to \citet{Jones1834}, the Shebli and Tabriz earthquakes had surface ruptures of more than 35 and 42 km, respectively; and the latter presented vertical separations of about 2 to 4 m. Other notable events include the $M$ 5.5 1807 Tasuj and the $M$ 6.7 1879 South Bozqush earthquakes. More recently, the same region was stroke by the $M_w$ 6.1 1997 Ardabil and $M_w$ 6.4 2012 Tabriz earthquakes, which caused extensive damage and took the lives of more than 1500 people \citep{USGS_ardabil, USGS_tabriz}.

Transitioning from the Azerbaijan to the Alborz seismic zone, the next fault system runs along the Rocks of the Talesh mountains, to the northwest of the Alborz Mountains. Faults in this area have ruptured recently \citep{Berberian1999}.
% , including an $M_s$ 6.0 low-angle thrust earthquake in 1978 \citep{Berberian1999} and 
The $M_w$ 7.4 1990 Manjil-Rudbar earthquake, for instance, caused numerous deaths and significant damage to the region in the south Caspian depression \citep{USGS_manjil}. At the central area of the Alborz seismic zone, near Tehran, the region is characterized by active reverse faults in the North Tehran Thrust (NTT) system. These faults are parallel to the northwest-trending structural gain of the Alborz Mountains belt, and include the south-dipping partially blind faults of Davudieh, Shian, and Bagh-E Feyz. According to \citet{Ambraseys2005}, Tehran---or the ancient city of Rey---has been devastated in multiple occasions due to severe earthquakes in this area. Historical accounts document the occurrence of multiple $M>7$ earthquakes in 743, 958, 1177, 1665, and 1830. The 1830 earthquake, in particular, ruptured the central segments of the NTT, adjacent to the Mosha fault.

To the east, the Kopeh-Dagh fault system is the predominant faulting system in the seismic zone of the same name. This system consists of several partly overlapping segments parallel to a NW--SE oriented structure of step-overs characterized by short south-dipping thrust faults striking E--W \citep{Berberian2001}, which exhibits active tectonic displacements along more than 500 km \citep{Trifonov1978}. This system is responsible for the $M_w$ 7.3 earthquake that stroke Ashgabat, the capital city of Turkmenistan and destroyed more than 30 villages in Iran in 1948. Historically, the Kopeh-Dagh seismic zone is also responsible for multiple earthquakes at the boundary between the Neyshabur and Binalud reverse faults between 1209 and 1405 \citep{Berberian1999}, and an estimated $M_s$ 7.1 earthquake in 10 A.D., roughly 30 kilometers from the border of Iran \citep{Berberian2001}. These and other notable earthquakes in the seismic zones of Azerbaijan and Alborz are shown in Fig.~\ref{fig:selected}.

@@

% With respect to different seismicity rates of the regions \citep{Nemati2015}, as well as the completeness of the seismic catalog \citep{Zare2014} , the study region is divided into three tectonic seismic zones. Each of these zones is characterized by its respective seismicity parameters values, including b-value and maximum magnitude ($M{_{max}}$). Part of Central Iran and Zagros regions are inside of the study region. We also calculated the seismicity parameters in these regions. Fig.~\ref{fig:Iran} shows the study region and different seismotectonic classification.

% \begin{figure*}[!ht] 
% \centering
% \includegraphics[scale=0.7]{figures/pdf/Figure1.pdf} 
% \caption{ a) Map of Iran and surrounding countries. The study area is presented in gray box. b) The study area containing seismotectonic provinces after \citet{Mirzaei1998}, and location of big cities. Dashed line is the subdivision that is proposed by \citet{Karimiparidari2013}. }

% \label{fig:Iran}
% \end{figure*}

% \subsection{Tectonic Regions and Seismic Faults}

% \subsubsection{Alborz Mountain Range (Alborz)}
% Tehran region has active reverse faults, which are parallel to the northwest-trending structural gain of the Alborz Mountains belt. Here a series of historical earthquakes occurred within a time period of more than 1100 years. Four earthquakes with magnitude more than 7 devastated the Tehran region in four centuries period, from 743 to 1177, but in the last 800 years, only one earthquake at 1830 has struck this region. At least three damaging earthquakes in 958 (western segments), 1655 (eastern segments), and 1830 (central segments), ruptured adjacent segments of the Mosha fault, located in northern Tehran.
% The North Tehran Thrust (NTT) adds more complexity due to the presence of south-dipping reverse faults, which are in part blind, such as the Davudieh, Shian, and Bagh-E Feyz.
% The northwest continuation of the Alborz Mountains, known as the Rocks of the Talesh Mountains, have been thrust northeastward and eastward over rocks of the south Caspian depression. An earthquake with Ms 6.0 in 1978 led to a focal mechanism consistent with a low-angle thrust \citep{Berberian1999}. Recently, the 1990 $M_w$ 7.4 Manjil-Rudbar earthquake caused extensive loss of life and significant damage \citep{USGS_manjil}. Based on existing documents, Tehran city and ancient Rey city have been destroyed completely several times by severe earthquakes with magnitudes greater than 7 \citep{Ambraseys2005}. \\

% \subsubsection{Azerbaijan}
% There were three earthquakes from 1721-86 that ruptured the North Tabriz Fault system from southeast to northwest. The Tabriz region is in the Araxes structural block of northwestern Iran, southwest of the continuation of the western Alborz Mountains towards the Caucasus. The North Tabriz Fault (NTF) is a complex northwest-trending structure, which contains evidence observed on aerial photographs, and vertical displacement with the north side up, of right-lateral strike-slip displacement \citep{Berberian1999}.
% The NTF system and nearby reverse faults ruptured from southeast to northwest in three earthquakes over 65 years: the Shebli earthquake with magnitude 7.3 in 1721 on the southeastern NTF with a surface rupture more than 35 km long, reported by \citet{Jones1834} the Tabriz earthquake with magnitude 7.4 in 1780 on the northwestern NTF, with a surface rupture more than 42 km long and vertical separation of 2 to 4 $m$; and the Marand-Mishu earthquake with magnitude of 6.3 in 1786 on the Mishu reverse fault and the Sufian segment of the NTF. Another earthquake with  magnitude 5.5 struck the Tasuj reverse fault farther west in 1807 and an earthquake of M 6.7 took place along the South Bozqush reverse fault farther southeast in 1879. Prior to the 1721-86 earthquake sequence, Tabriz was shaken by earthquakes in 858 (M 6.0), 1042 (M 7.3), 1273 (M ~ 6.5), and 1304 (M 6.7)\citep{Berberian1999}. Most recently, earthquakes with $M_w$ 6.1 in 1997 and $M_w$ 6.4 in 2012 occurred in Ardabil and Tabriz, respectively. These earthquakes caused around 1500 casualties and extensive damage \citep{USGS_ardabil,USGS_tabriz}.\\

% \subsubsection{Kopeh Dagh}
% The main Kopeh-Dagh fault system has experienced some historic earthquakes. Ashgabat, the capital city of Turkmenistan, was destroyed by an earthquake of Ms 7.2 in 1948 and destroyed more than 30 villages in Iran. This was the strongest earthquake to strike this region since at least 1455.
% The main Kopeh-Dagh fault consists of several partly overlapping segments parallel to the overall $NW - SE$ structure with step-overs. The regions of overlap are characterized by shorter south-dipping thrust faults striking about $E - W$ \citep{Berberian2001}. \citet{Trifonov1978} reported active displacement along the main Kopeh-Dagh fault for more than 500 km. 
% Massive destruction of the capital city of Mithradatkert is attributed to the 10 BC event Ms 7.1, roughly 30 kilometers from the border of Iran (Nesa mound) \citep{Berberian2001}.
% The Neyshabur sequence of four earthquakes between 1209 and 1405 respected the segment boundary between the Neyshabur and Binalud reverse fault system \citep{Berberian1999}.
% Historical records show that in 1209, the district of Neyshabur from Neyshabur city in the west to Daneh village in the east was totally destroyed \citep{Berberian1999}.\\

% \subsection{Historical  and Instrumental Seismicity}
% Uniform earthquake catalog is the most important factor in seismic hazard analysis of a region.  Different studies have been conducted in order to prepare a uniform catalog for Iranian Earthquake. Historical event interpretation and different magnitude conversion relationships are some of the factors which make all these catalogs different. Among different sources \citet{Ambraseys2005}, \citet{Berberian1994}, and \citet{moinfar1994} are some of the main sources for Iranian earthquake catalog. These catalogs contain historical and instrumental earthquakes that have been reported in literature and national and international networks. According to \citet{Ambraseys2005}, there are some scattered indication of earthquake effects  back to third millinuim BC. However, adequate documentary coverage of individual events begins from seventh century A.D. \citet{Mirzaei1997} provided a comprehensive list of studies that have been done in compilation a uniform catalog for Iranian earthquakes in order to prepare a uniform catalog of earthquake (all earthquakes are converted to surface wave magnitude Ms) for seismic hazard assessment in Iran. The catalog is covering the period of 4th century B.C. through 1994. The instrumental data is achievable from three national agencies including IIEES, International Institute of Earthquake Engineering and Seismology; IRSC, The Iranian seismological center, University of Tehran; and BHRC, Building and Housing Research Center (Iran Strong Motion Network) (For more information about these networks see \citet{Karimiparidari2013}).
% Many most of attenuation relationship use  moment magnitude $M_w$ as an input for the equation \citep{Douglas2011}. Moment magnitude is not suffering from saturation and has physical meaning \citep{Kanamori1977}. Due to mentioned reasons Mw has become a most appropriate magnitude scale in recent studies. 
%  \citet{Karimiparidari2013}compiled a uniform earthquake catalog for Iran and adjacent areas, using international and national databanks until April 2010.  They developed relationships between moment magnitude and other magnitude based on orthogonal regression.  They removed the dependent events (aftershocks and foreshocks) using the procedure by \citet{Gardner1974}. The catalog of events with magnitude equal and above Mw 5.5 is provided. 
% \citet{Shahvar2013} presented a unified and homogeneous catalog for the Iranian plateau $(M_w >= 4)$, created by merging data from two local catalogs and seven international agencies for 1900-2011 period. They used orthogonal regression method \citep{Castellaro2006} to derive magnitude conversion relation. By removing foreshocks and aftershocks according to the procedure detailed in Uhrhammer(1986), they also provided declustered version of the catalog.

>>>>

\begin{figure*}[ht!] 
	\centering
	\includegraphics[width=0.9\textwidth]{figures/pdf/figure-03}
	\caption{Seismic catalog. (a) Historical seismicity prior to 1900. (b) Instrumental seismicity. Size of symbols is proportional to the magnitude of the events. In the case of the historical seismicity catalog, the color fill indicates the year of the event.}
	\label{fig:selected}
\end{figure*}

\section{Seismic Catalog}

A fundamental step in regional seismic hazard analysis is the compilation of a reliable seismic catalog that includes both historically documented and instrumentally recorded earthquakes. Instrumental data are easily accessible through agencies such as the International Institute of Earthquake Engineering and Seismology (IIEES), the Iranian Seismological Center (IRSC) at the University of Tehran, or the Iranian Strong Motion Network operated by the Building and Housing Research Center (BHRC). Historical data, on the other hand, lends itself to subjective interpretation. Notable previous efforts to compose uniform earthquake catalogs for Iran include but are not limited to \citet{moinfar1994}, \citet{Berberian1994}, and \citet{Ambraseys2005}. These studies, however, differ from each other in their use of magnitude scales or magnitude conversion rules, and the extent to which they accept historical accounts as valid data. According to \citet{Ambraseys2005}, for instance, there is evidence of earthquakes that goes as far back as the third millennium B.C., yet some of those accounts are difficult to interpret. \citet{Mirzaei1997} compiled a catalog with historical events dating back to the 4th century B.C.~and instrumental earthquakes through 1994 in which a comprehensive list of studies was considered and magnitudes uniformly converted to surface wave magnitude ($M_s$).

Most ground motion prediction equations used today for seismic hazard assessments, however, use moment magnitude ($M_w$) instead, as it is well-known that this magnitude scale does not suffer from saturation issues as other scales, and provides a meaningful relationship with the physics of the rupture. Recent catalogs such as that by \citet{Karimiparidari2013}, \citet{Shahvar2013}, and \citet{Zare2014} are all compiled using magnitudes converted to $M_w$. \citet{Karimiparidari2013}, for instance, compiled a catalog for Iran and adjacent areas using international and national datasets including events up until April, 2010 with $M_w > 5$. They developed relationships between moment magnitude and other magnitude scales and removed dependent events (foreshocks and aftershocks) using the procedure by \citet{Gardner1974}. \citet{Shahvar2013} built a catalog for the Iranian plateau for events with $M_w \geq 4$ merging data from two local and seven international datasets for the 1900--2011 period. They used the orthogonal regression method by \citet{Castellaro2006} to derive magnitude conversions and removed foreshocks and aftershocks according to the procedure by \textcolor{red}{Uhrhammer (1986)}. In addition, \citet{Shahvar2013} provided a declustered version of the catalog.

In turn, the catalog by \citet{Zare2014} uses both historical data prior to 1900 and instrumental data up until 2006, and includes events from Afghanistan-Pakis\-tan, Saudi Arabia, the Caucasus, and parts of Turkey in addition to the Alborz-Azerbaijan, Zagros, Central Iran, Kopeh-Dagh and Makran seismic zones. All events in this catalog are converted to $M_w$ using relationships from previous studies \textcolor{red}{(e.g., pending ref.)}; and data were declustered using various methods including \citet{Gardner1974}, \citet{Reasenberg1985}, \textcolor{red}{Uhrhammer (1986), and Gruenthal (????) [is this level of detail really necessary?]}. In addition, \citet{Zare2014} provided the catalog completeness for different magnitude ranges using the cumulative frequency-magnitude distribution of \citet{Gutenberg1944} and \citet{richter1958}, and the frequency magnitude distribution obtained using ZMAP \citep{Wiemer2001}, and found the methods to yield consistent results. This catalog is particularly relevant because it is published as part of the current effort by the Global Earthquake Model (GEM) partnership and the Earthquake Model of the Middle East (EMME) project.

In this study we consider earthquakes with $M_w \geq 3$...

 % Earthquakes with magnitude 3 are not considered as a structural threat, however, the epicenter of earthquakes with magnitude $M_w 3$ are a  susceptible location for future bigger earthquakes.  According to our preliminary data processing based on IIEES data, Iranian catalog for earthquake with magnitude greater and equal 3, is complete from 2005. In order to make sure about the completeness of the catalog, we use IIEES data from 2000. We used the EMME project data provided by \citet{Zare2014} up to 2000. Fig.~\ref{fig:historical} represents the historical data (pre 1900) in the study region.

@@

% \subsection{Historical  and Instrumental Seismicity}

% Uniform earthquake catalog is the most important factor in seismic hazard analysis of a region.  Different studies have been conducted in order to prepare a uniform catalog for Iranian Earthquake. Historical event interpretation and different magnitude conversion relationships are some of the factors which make all these catalogs different. Among different sources \citet{Ambraseys2005}, \citet{Berberian1994}, and \citet{moinfar1994} are some of the main sources for Iranian earthquake catalog. These catalogs contain historical and instrumental earthquakes that have been reported in literature and national and international networks. According to \citet{Ambraseys2005}, there are some scattered indication of earthquake effects  back to third millinuim BC. However, adequate documentary coverage of individual events begins from seventh century A.D. \citet{Mirzaei1997} provided a comprehensive list of studies that have been done in compilation a uniform catalog for Iranian earthquakes in order to prepare a uniform catalog of earthquake (all earthquakes are converted to surface wave magnitude Ms) for seismic hazard assessment in Iran. The catalog is covering the period of 4th century B.C. through 1994. The instrumental data is achievable from three national agencies including IIEES, International Institute of Earthquake Engineering and Seismology; IRSC, The Iranian seismological center, University of Tehran; and BHRC, Building and Housing Research Center (Iran Strong Motion Network) (For more information about these networks see \citet{Karimiparidari2013}).

% Many most of attenuation relationship use  moment magnitude $M_w$ as an input for the equation \citep{Douglas2011}. Moment magnitude is not suffering from saturation and has physical meaning \citep{Kanamori1977}. Due to mentioned reasons Mw has become a most appropriate magnitude scale in recent studies.

%  \citet{Karimiparidari2013}compiled a uniform earthquake catalog for Iran and adjacent areas, using international and national databanks until April 2010.  They developed relationships between moment magnitude and other magnitude based on orthogonal regression.  They removed the dependent events (aftershocks and foreshocks) using the procedure by \citet{Gardner1974}. The catalog of events with magnitude equal and above Mw 5.5 is provided. 
% \citet{Shahvar2013} presented a unified and homogeneous catalog for the Iranian plateau $(M_w >= 4)$, created by merging data from two local catalogs and seven international agencies for 1900-2011 period. They used orthogonal regression method \citep{Castellaro2006} to derive magnitude conversion relation. By removing foreshocks and aftershocks according to the procedure detailed in Uhrhammer(1986), they also provided declustered version of the catalog.
